Type theories as quotient inductive-recursive types

### 2025-02-06 tex/main.tex
tried to polish abstract

\begin{abstract}
  \LT[noinline]{Try to interest Thorsten Altenkirch in the abstract}
  Defining type theory in type theory is a long-standing challenge and a benchmark for the expressiveness of a proof assistant.
  Type theory consisting of well-typed terms only is defined using quotient inductive-inductive type by Altenkirch and Kaposi, but well-typed terms are littered with explicit coercions, i.e.\ \transp's along equality constructors, and proofs need to account for coercions painstakingly alongside the interesting part.
  This mess has been dubbed the `transport hell' but typically suppressed in presentation for clarity, hiding the gap between the intention and the formalisation of type theory in type theory.
  In this paper, we aim for shortening the gap using quotient inductive-inductive-recursive types and definitions by overlapping patterns altogether, reducing the use of \transp's.
  As a case study, we investigate (parallel) substitution calculus and type theory with a universe and $\Pi$-types and compare ours with quotient inductive-inductive definitions.
\end{abstract}

>>>>

\begin{abstract}
  \LT[noinline]{Try to interest Thorsten Altenkirch in the abstract}
  Defining type theory in type theory is a long-standing challenge and a benchmark for the expressiveness of a proof assistant.
  Altenkirch and Kaposi used quotient inductive-inductive types to represent exactly the well-typed terms of type theory, with the price to be paid that terms frequently contain transports along equality constructors.
  As a result, proofs need to painstakingly account for coercions alongside the interesting part.
  This %mess
  has been dubbed `transport hell', but is typically suppressed in informal presentation for clarity, hiding the gap between the intention and the formalisation of type theory in type theory.
  In this paper, we aim to shorten this gap, using quotient inductive-inductive-recursive types and definitions by overlapping patterns, reducing the use of transports.
  As a case study, we investigate the (parallel) substitution calculus and type theory with a universe and $\Pi$-types, and compare ours with quotient inductive-inductive definitions.
\end{abstract}
